On The Impact of Atoms of Confusion in JavaScript Code

### 2020-11-16 3_Method.tex
updates na secao de MSR

To understand how often atoms of confusion appear in real settings, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written is JavaScript. We measured popularity using the project's stargazers, and selected only projects with stargazers\_count > 100. This metric, available on  GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs}.

After filtering out JavaScript project candidates, in the second step we built a curate dataset comprising the top 100 most popular repositories. Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}. Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research. The size of the projects range from small ones (883 liens of code) to complex systems with more than 10 MLOC. All projects in our dataset have at least \num{1161} forks and \num{23643} stars. We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.

In the third step we mined atoms of confusion from the repositories in our curate dataset. We written source code queries usng the CodeQL language~\cite{moor:gttse2007}, in order to filter the atoms of confusion. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as Java and JavaScript). 
We also automate the process of running the queries and exporting the results to a format that simplifies our analysis and the reproduction of this study. In the final step we carry out basic descriptive statistics, in order to understand how often the atoms of confusion appear in practice.

\begin{table*}[ht]
 \centering
 \begin{tabular}{rrrrrrr}
   \hline
             & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ \hline
 Lines of Code           & \num{883}  & \num{14939} & \num{39520} & \num{281088.46} & \num{145990.75} & \num{10149685} \\
 Num. of Contributors    & \num{6}   & \num{121} & \num{243} & \num{431.82} & \num{436.75} & \num{4042} \\
 Num. of Forks           & \num{1161}   & \num{3365} & \num{5711} & \num{8201.96} & \num{8438.25} & \num{68826} \\
 Num. of Stars        & \num{23643} & \num{27542.25} & \num{35687.5} & \num{45284.86} & \num{47963.25} & \num{310725} \\
    \hline
 \end{tabular}
 \caption{Some descriptive statistics about the projects used in the study}
 \label{tab:projects-statistics} 
 \end{table*}

>>>>

To understand how often atoms of confusion appear in real settings, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written is JavaScript. We measured popularity using the project's stargazers, and selected only projects with stargazers\_count greater than 100. This metric, available on  GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs}. We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.

After filtering out JavaScript project candidates, in the second step we built a curate dataset comprising the top \minedprojects repositories. Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}. Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research. The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC. All projects in our dataset have at least \num{1244} forks and \num{23672} stars. We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.

In the third step we mined atoms of confusion from the repositories in our curate dataset. We written source code queries using the CodeQL language~\cite{moor:gttse2007}, in order to filter the atoms of confusion. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as Java and JavaScript). 
We also automate the process of running the queries and exporting the results to a format that simplifies our analysis and the reproduction of this study. In the final step we carry out basic descriptive statistics, in order to understand how often the atoms of confusion appear in practice.

\begin{table*}[ht]
 \centering
 \begin{tabular}{rrrrrrr}
   \hline
             & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ \hline
 Lines of Code           & \num{5543}  & \num{16060.25} & \num{36161.5} & \num{111432.33} & \num{111182.50} & \num{1278405} \\
 Num. of Forks     & \num{1244}   & \num{3029.75} & \num{6078} & \num{8906.24} & \num{9260.50} & \num{68849} \\
 Num. of Contributors  & \num{6}   & \num{184} & \num{285.5} & \num{533.44} & \num{515.75} & \num{4047} \\
 Num. of Stars        & \num{23672} & \num{27055} & \num{34990} & \num{46919.51} & \num{48518} & \num{310935} \\

    \hline
 \end{tabular}
 \caption{Some descriptive statistics about the projects used in the study}
 \label{tab:projects-statistics} 
 \end{table*}

### 2020-11-16 4_Result.tex
updates na secao de MSR

We mined \minedprojects open source JavaScript repositories to understand how often atoms of confusion arise in real software. Similarly to a previous study~\cite{DBLP:conf/msr/GopsteinZFC18}, which investigates the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion frequently arise in JavaScript open source systems.

The 3 most frequently found atoms occur in at least 80\% of the projects.
Considering the extremes, atom Ternary Operator was found in 89.69\% of the repositories, while atom Comma Operator occurred in only 11.34\% of them (as seen in Figure~\ref{fig:rate}).

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{images/rate-1.pdf}
    \caption{Incidence of atoms of confusion across the \minedprojects JavaScript repositories we mined}
    \label{fig:rate}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{images/chart_ocurrence_kloc.png}
    \caption{The rate of each atom per thousand lines of code (KLOC) across all \minedprojects repositories}
    \label{fig:atoms-occurrence}
\end{figure}

Considering all JavaScript projects in our dataset, we found 
a total of \num{206317} atoms of confusion, though three atoms are responsible for 87.47\% of this total: Post Increment Expression, Omitted Curly Braces, and the Ternary Operator. The remaining atoms accounts for \num{25851} (12.53\%) of the total number of occurrences, and the Comma Operator and and Arithmetic as Control Flow are the atoms that arise less frequently (0.25\% in total with 373 and 164 occurrences, respectively).
%In our dataset of JavaScript projects, 
We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}.

The results of our survey and interviews suggest that the Ternary Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey). Nonetheless, the Post Increment Expression and Omitted Curly Braces atoms are listed in the top three sources of misunderstanding 
(see Table~\ref{tab:difference-correctness}). As such, fixing the the Post Increment Expression and Omitted Curly Braces might solve most of the unclear code caused by atoms of confusion.

>>>>

We mined \minedprojects open source JavaScript repositories to understand how often atoms of confusion arise in real software. Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18}~\cite{DBLP:journals/ese/MedeirosLAAKRG19}, which investigates the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion frequently arise in JavaScript open source systems.

The 5 most frequently found atoms occur in at least 80\% of the projects.
Considering the extremes, atoms Ternary Operator and Access Property were found in all repositories, while atom Comma Operator occurred only in 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).

@@

% latex table generated in R 4.0.3 by xtable 1.8-4 package
% Fri Nov 13 11:28:56 2020
\begin{table}[!htb]
\centering
 \caption{Summary of atoms occurrences on our dataset}

\setlength\tabcolsep{2pt} % default value: 6pt
\label{tab:occurrences-summary}
\begin{tabular}{lrcc}

  \hline
Atom & Projects & Ocurrences/KLOC & Category \\ 
 \hline
Ternary Operator & 100\% & 10.16 &  highly used \\ 
  Access Property & 100\% & 3.63 &  highly used \\ 
  Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
  Post Increment & 90.28\% & 5.62 &  highly used \\ 
  Arrow Function & 83.33\% & 21.10 &  highly used \\ 
  Change Literal Encoding & 76.39\% & 21.43 &  highly used \\ 
  Object Destructing & 72.22\% & 2.86 &  highly used \\ 
  Pre-Increment & 83.33\% & 1.04 & commonly used \\ 
  Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
  Implicit Operand Conversion & 79.17\% & 0.80 & commonly used \\ 
  Array Destructing & 62.50\% & 0.47 & commonly used \\ 
  Array Spread & 59.72\% & 0.31 & commonly used \\ 
  Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
  Object Spread & 55.56\% & 0.64 & commonly used \\ 
  Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
  Comma Operator & 15.28\% & 0.03 & little used \\ 
   \hline
\end{tabular}
\end{table}

Considering all JavaScript projects in our dataset, we found 
a total of \num{615364} atoms of confusion, though five atoms are responsible for 84.65\% of this total: Arrow Function, Change Literal Encoding, Ternary Operator, Omitted Curly Braces and the Post Increment. The remaining atoms accounts for \num{94450} (15.35\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.0064\% in total with 232 and 165 occurrences, respectively).

Regarding frequency, we proceeded with a classification of atom occurrences according to the metrics established by a previous study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. As such, we found 7 atoms as highly used (43.75\%), 7 as commonly used (43.75\%) and only 2 atoms as little used (12.50\%). We also calculated the frequency of each atom per KLOC and found atoms Change Literal Encoding and Arrow Function as the most occurring in practice (21.43 and 21.10 occurrences per KLOC, respectively).
%In our dataset of JavaScript projects, 
%We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}. %

The results of our survey and interviews suggest that the Ternary Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey). Nonetheless, the Post Increment Expression and Omitted Curly Braces atoms are listed in the top three sources of misunderstanding 
(see Table~\ref{tab:difference-correctness}). As such, fixing the the Post Increment and Omitted Curly Braces might solve most of the unclear code caused by atoms of confusion.
